### 2020-12-19 paper/sex_selection_spacing-ver5.tex
Add measure of education paragraph to Data

I restrict the sample to Hindus for two reasons.
First, Hindus are the majority population group, about 80\% of India’s population.
Second, the prior literature shows that son preference and use of sex selection vary 
substantially between Hindus and the second largest group, Muslims. 
Combining them and assuming that the baseline hazard is the same would lead to biased 
results.
Because of space constraints and the relatively small number of observations once split 
by education and periods, I do not provide separate results for Muslims or any of the 
remaining groups, such as Sikhs, Jains, and Christians.

@@

In addition to a large number of women surveyed and the long period covered, a significant 
benefit of the NFHS over other surveys is that enumerators pay careful attention to the 
spacing between births and probe for ``missed'' births.
For India, the main concern is the underreporting of deceased children, especially a systematic 
recall error where respondents' likelihood of reporting the birth of a deceased child depends 
on the sex of that child.
Unreported deceased children inflate the length of birth intervals and, with declining mortality, 
make changes over time appear too small.
In the online appendix, I provide a detailed analysis of systematic recall error, which 
shows that recall error depends heavily on how long ago a woman was married. 
I, consequently, drop women married 22 years or more.%
\footnote{
Recall error is likely behind the designation of the first two rounds of NFHS as ``moderate 
quality'' in an analysis of the quality of birth histories in DHS surveys and its impact 
on fertility estimates \citep{Schoumaker2014}.
}

@@

Direct information on the use of sex selection is not available, so I compare periods 
based on the changes in access and legality of prenatal sex determination in India.
Abortion has been legal in India since 1971.
Reports of sex determination appeared around 1982--1983, and the number of clinics 
quickly increased \citep{Sudha1999,bhat06,Grover2006}.
In 1994, the Prenatal Diagnostic Techniques Act made determining and communicating 
the sex of a fetus illegal.%
\footnote{
% Details about the act are at \href{http://pndt.gov.in/}{http://pndt.gov.in/}.
There is little evidence that the ban significantly affected sex ratios \citep{Das-Gupta2019}.
}
Finally, although sex selection increased even after 1994, we may have passed a 
turning point in its use in the mid-2000s \citep{Das_Gupta2009,Kumar2012,Bongaarts2013,Diamond-Smith2015}.

@@

The allocation of spells into periods is determined by when conception, and, therefore, 
decisions on sex selection can begin. 
Hence, some spells cover two periods, which may bias downward the differences between the 
periods.
Most sterilizations take place soon after giving birth. 
These spells, therefore, do not show up in the samples used. 
Furthermore, sterilization depends strongly on the sex composition of prior children
with lower probabilities, the fewer boys. 
The effect is to bias downward the differences in parity progression probabilities.

I divide the explanatory variables into two groups, nonproportional and proportional. 
The first group consists of characteristics shown in the prior literature to affect the 
spacing choice and the use of sex selection: Mother’s education, sex composition of 
previous children, and area of residence. 
To minimize any potential bias from including proportional variables, I estimate a 
separate model for each birth interval, education group, and period combination, rather 
than including education as a variable. 
I capture sex composition with dummy variables for the possible combinations, ignoring 
the ordering of births.%
\footnote{
With sex selection, the composition of prior children is, in principle, endogenous. 
It is beyond the scope of this paper to develop a method for dealing with this issue. 
}
Area of residence is a dummy variable for living in an urban area.

The second group of variables consists of those expected to have an approximately 
proportional effect on the hazard.
These include the mother's age when the spell begins, the household's
land ownership, and whether it belongs to a scheduled tribe or caste.
Appendix Tables \ref{tab:des_stat1} and \ref{tab:des_stat2} present descriptive statistics.

>>>>

I restrict the sample to Hindus for two reasons.
First, Hindus are the majority population group, about 80\% of India’s population.
Second, the prior literature shows that son preference and sex selection vary 
substantially between Hindus and the second largest group, Muslims. 
Combining them and assuming that the baseline hazard is the same would lead to biased 
results.
Because of space constraints and the relatively small number of observations once split 
by education and periods, I do not provide separate results for Muslims or any  
remaining groups, such as Sikhs, Jains, and Christians.

@@

In addition to the large number of women surveyed and the long period covered, a significant 
benefit of the NFHS over other surveys is that enumerators pay careful attention to the 
spacing between births and probe for ``missed'' births.
For India, the main concern is the underreporting of deceased children, especially a systematic 
recall error where respondents' likelihood of reporting the birth of a deceased child depends 
on the sex of that child.
Unreported deceased children inflate the length of birth intervals and, with declining mortality, 
make changes over time appear too small.
In the online appendix, I provide a detailed analysis of systematic recall error, which 
shows that recall error depends heavily on how long ago a woman was married. 
I, consequently, drop women married 22 years or more.%
\footnote{
Recall error is likely behind the designation of the first two rounds of NFHS as ``moderate 
quality'' in an analysis of the quality of birth histories in DHS surveys and its impact 
on fertility estimates \citep{Schoumaker2014}.
}

@@

Direct information on sex selection is not available, so I compare periods 
based on the changes in access and legality of prenatal sex determination in India.
Abortion has been legal in India since 1971.
Reports of sex determination appeared around 1982--1983, and the number of clinics 
quickly increased \citep{Sudha1999,bhat06,Grover2006}.
In 1994, the Prenatal Diagnostic Techniques Act made determining and communicating 
the sex of a fetus illegal.%
\footnote{
% Details about the act are at \href{http://pndt.gov.in/}{http://pndt.gov.in/}.
There is little evidence that the ban significantly affected sex ratios \citep{Das-Gupta2019}.
}
Finally, although sex selection increased even after 1994, we may have passed a 
turning point in its use in the mid-2000s \citep{Das_Gupta2009,Kumar2012,Bongaarts2013,Diamond-Smith2015}.

@@

The allocation of spells into periods is determined by when conception, and, therefore, 
decisions on sex selection can begin. 
Hence, some spells cover two periods, which may bias downward the differences between the 
periods.
Most sterilizations take place soon after giving birth. 
These spells, therefore, do not show up in the samples used. 
Furthermore, sterilization depends strongly on prior children's sex composition; the 
fewer boys, the lower the probability of sterilization. 
The effect is to bias downward the differences in parity progression probabilities.

I divide the explanatory variables into two groups, nonproportional and proportional. 
The first group consists of characteristics shown in the prior literature to affect the 
spacing choice and the use of sex selection: Mother’s education, sex composition of 
previous children, and area of residence.
To minimize any potential bias from including proportional variables, I estimate a 
separate model for each birth interval, education group, and period combination, rather 
than including education as a variable.

I divide education levels into four groups: No education, 1--7 years, 8--11 years, and 
12 and more years.
The latter two correspond to having completed primary and secondary school, 
respectively.%
\footnote{
Although there are variations by state, elementary education in India consists of a 
primary school covering grades one through five and an upper primary---or middle 
school---covering grades six through eight. 
Similarly, secondary education covers grades nine and tenth for ``secondary education'' 
and 11 and 12 for ``upper secondary.''
}
To ensure that the results are comparable with the prior literature on fertility
and mortality in India, I follow the NFHS reports, except that I combine the 
less than five years and 5--7 years of schooling completed and the 8--9 
and 10--11 years of schooling completed to ensure sufficient cell sizes.

I capture sex composition with dummy variables for the possible combinations, ignoring 
the ordering of births.%
\footnote{
With sex selection, the composition of prior children is, in principle, endogenous. 
It is beyond the scope of this paper to develop a method for dealing with this issue. 
}
Area of residence is a dummy variable for living in an urban area.

The second group of variables consists of those expected to have an approximately 
proportional effect on the hazard.
These include the mother's age when the spell begins, the household's
land ownership, and whether it belongs to a scheduled tribe or caste.
Appendix Tables \ref{tab:des_stat1} and \ref{tab:des_stat2} present the descriptive 
statistics.
